Spin canonical divisors of log stacky curves

### 2015-06-25 stacky_write_up/spin_cring.tex
some intro changes (parts -> background), ref fix

\section{Introduction}
Let $\Gamma$ be a {\bf Fuchsian} subgroup, i.e. a discrete subgroup
of $PSL_2(\BR)$ acting on the {\bf upper half plane} $\bh$ by
fractional linear transformations. We can then consider the
orbifold $\Gamma \backslash \bh$ over $\BC$. For example, in the
case $\Gamma$ acts freely on $\bh$, $\Gamma \backslash \bh$ is a
Riemann Surface over $\BC$. It may well be that $\Gamma \backslash
\bh$ is not compact. Defining $\Delta$ as the divisor of cusps
associated to $\Gamma$, we can form a space $\Gamma \backslash
\bh^*$ which is a compactification of $\Gamma \backslash \bh,$ with
associated {\bf divisor of cusps} $\Delta$.

We can now consider the
{\bf ring of modular forms} $M(\Gamma) = \oplus_k M_k(\Gamma)$. In
this paper, for general $\Gamma$, we sharply bound the degree of
the generators and relations for $M(\Gamma).$ This was already
answered in \cite[Chapters 7-9]{vzb:stacky} in the case that
$M(\Gamma)$ has $M_{2k}(\Gamma) = 0$ for all $k \in \BZ$, that is,
when $\Gamma$ has no odd weight modular forms, and we extend this
to all Fuchsian subgroups $\Gamma$. In order to answer this
question, we translate the seemingly analytic question into the
algebraic category, using a generalization of the GAGA principle.
That is, by \cite[Proposition 6.1.5]{vzb:stacky}, there is an
equivalence of categories between orbifold curves and log stacky
curves over $\BC$. Hence, for the remainder of the paper, we will
work in the algebraic category. Let $X$ be a smooth algebraic
curve over $k$ of genus $g$. It is well known that the sheaf $\Omega
_X,$ with associated divisor $K_X$ determines the {\bf canonical map
} $\pi:X \rightarrow \BP_k^{g - 1}.$ Then, the {\bf canonical ring},
is defined as
\begin{align*}
	R(X,K_X) = \bigoplus_{n \geq 0} H^0(X,nK_X),
\end{align*}

\noindent
with multiplication structure corresponding to tensor product of
sections. Since $\Omega_X$ is ample, $X \cong \proj R$. Petri's
theorem gives generators and relations for $R(X,K_X),$ see for
example, \cite{saint-donat:proj}[Theorem, p. 157], showing that in
most cases $R(X,K_X)$ is generated in degree 1 with relations in
degree 2. This has the pleasant geometric consequence that all such
canonically embedded curves are scheme theoretically cut out by
quadratic equations.

Following Voight and Zureick-Brown \cite{vzb:stacky}, we examine a
much more general setting, generalizing Petri's theorem in many
directions. A divisor $\Delta \in d X$ is a {\bf log divisor} if
$\Delta = \sum_{i}^{}P_i \in \di X,$ with all $P_i$ distinct. Let
$(\sx,\Delta)$ denote a tame separably rooted curve over a
field $k$ so that $X$ is the coarse space of $\sx$, as defined in
\cite[Section 5.2]{vzb:stacky} so that $\di \sx = \di X \bigoplus_{i
= 1}^r \langle \frac{1}{e_i}P_i \rangle,$ with $e_i \in \BZ$. Let $K_
\sx = K_X + \sum_{i = 1}^{r}\frac{e_i - 1}{e_i}P_i.$ In this case, we
say $\sx$ has {\bf stabilizer order} $e_i$ at $P_i$. Call $\halfcan
\in \di \sx$ a {\bf log half-canonical divisor} on $(\sx,\Delta)$,
be a divisor $\halfcan \in \di \sx$ so that $2\halfcan \sim K_\sx +
\Delta.$ If $X$ has genus $g$ and $\delta = \deg \Delta,$ then we
say $\sx$ has signature $(g; e_1, \ldots, e_r;\delta)$. A {\bf log
spin curve} is a triple $(X,\Delta,\halfcan)$ where $(X,\Delta)$
and $\halfcan$ is a log half-canonical divisor on $(\sx,\Delta)$.
The central object of study in this paper is the {\bf log spin
canonical ring} of $(X,\Delta,L)$ is
\begin{align*}
	R(X,\Delta,L) = \bigoplus_{n \geq 0} H^0(X,nL)
\end{align*}

@@

\noindent
is generated as a $k$-algebra by elements of degree at most $e =
\max(5, e_1, \ldots, e_r).$ If the genus $g \geq 2,$ and $\Delta = 0$ then 
$R(\sx,\Delta, \halfcan)$ has minimal relations in degree at most $\max(11, 2e_1,
\ldots, 2e_r).$ Otherwise, $R(\sx, \Delta, \halfcan)$ has minimal relations in degree at most
$\max(10, 2e_1, \ldots, 2e_r),$ so long as $\sigma$ does not lie in a finite list of
exceptional cases, as listed in Table ~\ref{table:g-1-exceptional}
for signatures with $g = 1$ and Table ~\ref{table:g_0_exceptional}
for signatures with $g = 0$.
\end{thm}

@@

@@

\section{Inductive Lemmata}
\label{sec:induction}
First we present the general inductive Lemmata that will be applied
for all genera. We offer various inductive Lemmata that will allow
us to add fractional point $\frac{\alpha}{\beta}P$ to a $\BQ$
 divisor $D$ of $X$ and describe the new generators and relations
of $R_{D+p}$ not in $R_D$. Next, we provide a method specific to
the case of stacky canonical divisors, which
under certain conditions allows us to transfer information about
the generators and relations of the half canonical ring of a stacky
curve $\sx$' with signature $(g; e_1', \ldots, e_\tau'; 0)$ to a
stacky curve $\sx$ with signature $(g; e_1, \ldots, e_\tau; 0)$
where each $e_i\in \{e_i', e_i' + 2\}$.

@@

\todo{Include one of the following, if the first one, include a 
remark for quotient}
\iffalse{
\begin{lem}
\label{lem:minimal_quadratic}
Suppose $k[x_1, \ldots, x_n]$ is a graded ring, not necessarily
generated in degree 1, is equipped with a monomial ordering $\prec$.
 Let $\phi:k[x_1, \ldots, x_n] \rightarrow B'$ be a map of graded
rings with kernel $I'$, such that $I'$ is generated by elements $f_1,
\ldots, f_j$, where $in_\prec(f_i) = x_i x_j,$ and no term of $f_i$
is of the form $x_k$. Then, the elements $f_i$ determine minimal
generators for $I$.
\end{lem}
\begin{proof}
Regrade the ring $k[x_1, \ldots, x_n]$ so that all $x_i$ lie in
degree 1. Then,
\[
	\phi : k[x_1, \ldots, x_n] \rightarrow B
\]
\noindent
determines a map of rings, no longer necessarily graded. Consider
the grlex monomial ordering $\prec$ on $k[x_1, \ldots, x_n]$.
\todo{This ordering isn't quite right. 
We need to preserve the ordering on quadratic terms}
By
assumption, the kernel $I$ is generated by elements whose initial
terms have degree 2 and are all distinct. Therefore, $\dim_k (k[x_1,
\ldots, x_n])_2 = \binom{n}{2} - j.$ This implies any set of
generators for $I$ must have at least $j$ elements. Since $f_1,
\ldots, f_j$ is such a set, it is minimal.
\end{proof}

\begin{rem}
In particular, in the case of
\[
	\begin{tikzcd}
		I\ar[hookrightarrow]{r}\ar[hookrightarrow]{d} & k[x_1, \ldots, x_n]
		\ar[twoheadrightarrow]{r} \ar[hookrightarrow]{d} & B \ar[hookrightarrow]{d} \\
		I' \ar[hookrightarrow]{r} & k[x_1, \ldots, x_n, y_1, \ldots, y_m] \ar[twoheadrightarrow]{r}& B'
	\end{tikzcd}
\]
inducing
\[
	I/I'\hookrightarrow k[y_1, \ldots, y_m]\twoheadrightarrow B'/B
\]
\todo{how do we make this proof?}
\end{rem}
}\fi

\begin{lem}
\label{lem:minimal_quadratic_2}
Suppose $k[x_1, \ldots, x_n]\hookrightarrow k[x_1, \ldots, x_n, y_1, \ldots, y_m]$ be graded ring, not necessarily
generated in degree 1 that are equipped with monomial ordering $\prec$.
Let $\phi: k[x_1, \ldots, x_n] \twoheadrightarrow B, \phi':k[x_1, \ldots, x_n, y_1, \ldots, y_m]\twoheadrightarrow B'$ be maps of graded
rings with kernels $I \hookrightarrow I'$ such that the following diagram commutes:
\[
	\begin{tikzcd}
	I\ar[hookrightarrow]{r}\ar[hookrightarrow]{d} & k[x_1, \ldots, x_n]\ar[twoheadrightarrow]{r} \ar[hookrightarrow]{d} & B \ar[hookrightarrow]{d}\\
	I' \ar[hookrightarrow]{r} & k[x_1, \ldots, x_n, y_1, \ldots, y_m] \ar[twoheadrightarrow]{r}& B'
	\end{tikzcd}
\]
Further suppose that $I'$ is generated over $I$ by elements $f_1,
\ldots, f_j$ as a $k$-vector space, where $in_\prec(f_i) \in \{y_i x_j, y_i y_j\}$ and no monomial of $f_i$
is of the form $y_i$. Then, the elements $f_i$ determine minimal
generators for $I'$ over $I$.
\end{lem}

@@

There are two main cases: when $\delta \geq 2$ and when $\delta = 0$
 (note that $\delta$ must be odd or zero because $\halfcan \in \text
{ Div } \sx$). In the first case, $\deg \lfloor \halfcan \rfloor
\geq 0$ so the result directly follows from Lemma
~\ref{lem:deg1-sat-ind}. When $\delta = 0$, this is not necessarily
true so the proof is a rather involved induction using Lemma
~\ref{lem:sat-three-induction-g-0} and Theorem ~\ref{thm:ram_order_ind}.
The remainder of this section is dedicated to characterizing the
saturations (Subsection ~\ref{ssec:g_0_saturation}), inductive
base cases (Subsection ~\ref{ssec:g_0_base}), and exceptional
cases (Subsection ~\ref{ssec:g_0_exceptional}) when $\delta = 0$
before proving the main theorem in full for all $\delta$ in
Subsection ~\ref{ssec:g_0_main}.

@@

\ssec{Exceptional Cases}
\label{ssec:g_0_exceptional}
First, we describe the cases that would not be covered by induction.
Here we present the explicit generators and relations of the remaining
cases given by signatures in the finite set

@@

\noindent
is generated as a $k$-algebra by elements of degree at most $e =
\max(5, e_1, \ldots, e_r)$ and has relations in degree at most $2e$,
so long as $\sigma$ does not lie in a finite list of exceptional
cases, as listed in Table ~\ref{table:g_0_exceptional}.
\end{thm}

@@

Now suppose $\delta = 0$, so $\halfcan$ is not necessarily effective.
Let $\sigma = (0; e_1, \ldots, e_r; 0)$ be a signature with $r > 0$
and not one of the exceptional cases in Table
~\ref{table:g_0_exceptional}.

>>>>

\section{Introduction}
Let $\Gamma$ be a {\bf Fuchsian} subgroup, i.e. a discrete subgroup
of $PSL_2(\BR)$ acting on the upper half plane $\bh$ by fractional
linear transformations. We consider the {\bf ring of modular forms}
$M(\Gamma) = \bigoplus_k M_k(\Gamma)$. In this paper, for general
$\Gamma$, we sharply bound the degree of the generators and
relations for $M(\Gamma).$ The case where $M(\Gamma)$ has $M_{2k + 1
}(\Gamma) = 0$ for all $k \in \BZ$, i.e. when $\Gamma$ has no odd
weight modular forms, is answered in Voight and Zureick-Brown
\cite[Chapters 7-9]{vzb:stacky}. This work extends this result to
all Fuchsian subgroups $\Gamma$. In order to answer this question,
we translate the seemingly analytic question into the algebraic
category, using a generalization of the GAGA principle. By
\cite[Proposition 6.1.5]{vzb:stacky}, there is an equivalence of
categories between orbifold curves and log stacky curves over $\BC$
. Hence, for the remainder of the paper, we will work in the
algebraic category.

Let $X$ be a smooth algebraic curve over $k$ of genus $g$. It is
well known that the sheaf $\Omega _X,$ with associated divisor $K_X$
 determines the {\bf canonical map } $\pi:X \rightarrow \BP_k^{g - 1}$.
Then, the {\bf canonical ring} is defined to be
\begin{align*}
	R(X,K_X) = \bigoplus_{n \geq 0} H^0(X,nK_X),
\end{align*}

\noindent
with multiplication structure corresponding to tensor product of
sections. Since $\Omega_X$ is ample, $X \cong \proj R$. Petri's
theorem gives generators and relations for $R(X, K_X)$ showing that
in most cases $R(X, K_X)$ is generated in degree 1 with relations
in degree 2 (see \cite[p. 157]{saint-donat:proj}). This has the
pleasant geometric consequence that all such canonically embedded
curves are scheme theoretically cut out by quadratic equations

Following Voight and Zureick-Brown \cite{vzb:stacky}, we examine a
much more general setting, generalizing Petri's theorem in many
directions. A divisor $\Delta \in d X$ is a {\bf log divisor} if
$\Delta = \sum_{i}^{}P_i \in \di X,$ with all $P_i$ distinct. Let
$(\sx,\Delta)$ denote a tame separably rooted curve over a
field $k$ so that $X$ is the coarse space of $\sx$, as defined in
\cite[Section 5.2]{vzb:stacky} so that $\di \sx = \di X \bigoplus_{i
= 1}^r \langle \frac{1}{e_i}P_i \rangle,$ with $e_i \in \BZ$. Let $K_
\sx = K_X + \sum_{i = 1}^{r}\frac{e_i - 1}{e_i}P_i.$ In this case, we
say $\sx$ has {\bf stabilizer order} $e_i$ at $P_i$. Call $\halfcan
\in \di \sx$ a {\bf log half-canonical divisor} on $(\sx,\Delta)$,
be a divisor $\halfcan \in \di \sx$ so that $2\halfcan \sim K_\sx +
\Delta.$ If $X$ has genus $g$ and $\delta = \deg \Delta,$ then we
say $\sx$ has \textbf{signature} $(g; e_1, \ldots, e_r;\delta)$. A
{\bf log spin curve} is a triple $(X,\Delta,\halfcan)$ where $(X,\Delta)$
and $\halfcan$ is a log half-canonical divisor on $(\sx,\Delta)$.
The central object of study in this paper is the {\bf log spin
canonical ring} of $(X,\Delta,L)$ is
\begin{align*}
	R(X,\Delta,L) = \bigoplus_{n \geq 0} H^0(X,nL)
\end{align*}

@@

\noindent
is generated as a $k$-algebra by elements of degree at most $e =
\max(5, e_1, \ldots, e_r).$ If the genus $g \geq 2,$ and $\Delta = 0
$ then $R(\sx,\Delta, \halfcan)$ has relations generated in degrees
up to $\max(11, 2e_1, \ldots, 2e_r)$. Otherwise, $R(\sx, \Delta,
\halfcan)$ has relations generated in degrees up to $\max(10, 2e_1,
\ldots, 2e_r),$ so long as $\sigma$ does not lie in a finite list
of exceptional cases, as listed in Table ~\ref{table:g-1-exceptional}
for signatures with $g = 1$ and Table ~\ref{table:g-0-exceptional}
for signatures with $g = 0$.
\end{thm}

@@

\subsection{Junk}
This is stuff that has been moved out of the introduction to possibly
be retooled for this section.

We can then consider the
orbifold $\Gamma \backslash \bh$ over $\BC$. For example, in the
case $\Gamma$ acts freely on $\bh$, $\Gamma \backslash \bh$ is a
Riemann Surface over $\BC$. It may well be that $\Gamma \backslash
\bh$ is not compact. Defining $\Delta$ as the divisor of cusps
associated to $\Gamma$, we can form a space $\Gamma \backslash
\bh^*$ which is a compactification of $\Gamma \backslash \bh,$ with
associated {\bf divisor of cusps} $\Delta$.

@@

\section{Inductive Lemmata}
\label{sec:induction}
First we present the general inductive Lemmata that will be applied
for all genera. We offer various inductive Lemmata that will allow
us to add fractional point $\frac{\alpha}{\beta}P$ to a $\BQ$
 divisor $D$ of $X$ and describe the new generators and relations
of $R_{D + P}$ not in $R_D$. Next, we provide a method specific to
the case of stacky canonical divisors, which
under certain conditions allows us to transfer information about
the generators and relations of the half canonical ring of a stacky
curve $\sx$' with signature $(g; e_1', \ldots, e_\tau'; 0)$ to a
stacky curve $\sx$ with signature $(g; e_1, \ldots, e_\tau; 0)$
where each $e_i\in \{e_i', e_i' + 2\}$.

@@

\todo{Include one of the following, if the first one, include a 
remark for quotient}

%\begin{lem}
%\label{lem:minimal_quadratic}
%Suppose $k[x_1, \ldots, x_n]$ is a graded ring, not necessarily
%generated in degree 1, is equipped with a monomial ordering $\prec$.
 %Let $\phi:k[x_1, \ldots, x_n] \rightarrow B'$ be a map of graded
%rings with kernel $I'$, such that $I'$ is generated by elements $f_1,
%\ldots, f_j$, where $in_\prec(f_i) = x_i x_j,$ and no term of $f_i$
%is of the form $x_k$. Then, the elements $f_i$ determine minimal
%generators for $I$.
%\end{lem}
%\begin{proof}
%Regrade the ring $k[x_1, \ldots, x_n]$ so that all $x_i$ lie in
%degree 1. Then,
%\[
	%\phi : k[x_1, \ldots, x_n] \rightarrow B
%\]
%\noindent
%determines a map of rings, no longer necessarily graded. Consider
%the grlex monomial ordering $\prec$ on $k[x_1, \ldots, x_n]$.
%\todo{This ordering isn't quite right. 
%We need to preserve the ordering on quadratic terms}
%By
%assumption, the kernel $I$ is generated by elements whose initial
%terms have degree 2 and are all distinct. Therefore, $\dim_k (k[x_1,
%\ldots, x_n])_2 = \binom{n}{2} - j.$ This implies any set of
%generators for $I$ must have at least $j$ elements. Since $f_1,
%\ldots, f_j$ is such a set, it is minimal.
%\end{proof}
%
%\begin{rem}
%In particular, in the case of
%\[
	%\begin{tikzcd}
		%I\ar[hookrightarrow]{r}\ar[hookrightarrow]{d} & k[x_1, \ldots, x_n]
		%\ar[twoheadrightarrow]{r} \ar[hookrightarrow]{d} & B \ar[hookrightarrow]{d} \\
		%I' \ar[hookrightarrow]{r} & k[x_1, \ldots, x_n, y_1, \ldots, y_m] \ar[twoheadrightarrow]{r}& B'
	%\end{tikzcd}
%\]
%inducing
%\[
	%I/I'\hookrightarrow k[y_1, \ldots, y_m]\twoheadrightarrow B'/B
%\]
%\todo{how do we make this proof?}
%\end{rem}

\begin{lem}
\label{lem:minimal_quadratic}
Suppose $k[x_1, \ldots, x_n]\hookrightarrow k[x_1, \ldots, x_n, y_1, \ldots, y_m]$ be graded ring, not necessarily
generated in degree 1 that are equipped with monomial ordering $\prec$.
Let $\phi: k[x_1, \ldots, x_n] \twoheadrightarrow B, \phi':k[x_1, \ldots, x_n, y_1, \ldots, y_m]\twoheadrightarrow B'$ be maps of graded
rings with kernels $I \hookrightarrow I'$ such that the following diagram commutes:
\[
	\begin{tikzcd}
	I\ar[hookrightarrow]{r}\ar[hookrightarrow]{d} & k[x_1, \ldots, x_n]\ar[twoheadrightarrow]{r} \ar[hookrightarrow]{d} & B \ar[hookrightarrow]{d}\\
	I' \ar[hookrightarrow]{r} & k[x_1, \ldots, x_n, y_1, \ldots, y_m] \ar[twoheadrightarrow]{r}& B'
	\end{tikzcd}
\]
Further suppose that $I'$ is generated over $I$ by elements $f_1,
\ldots, f_j$ as a $k$-vector space, where $in_\prec(f_i) \in \{y_i x_j, y_i y_j\}$ and no monomial of $f_i$
is of the form $y_i$. Then, the elements $f_i$ determine minimal
generators for $I'$ over $I$.
\end{lem}

@@

There are two main cases: when $\delta \geq 2$ and when $\delta = 0$
 (note that $\delta$ must be odd or zero because $\halfcan \in \text
{ Div } \sx$). In the first case, $\deg \lfloor \halfcan \rfloor
\geq 0$ so the result directly follows from Lemma
~\ref{lem:deg1-sat-ind}. When $\delta = 0$, this is not necessarily
true so the proof is a rather involved induction using Lemma
~\ref{lem:sat-three-induction-g-0} and Theorem ~\ref{thm:ram_order_ind}.
The remainder of this section is dedicated to characterizing the
saturations (Subsection ~\ref{ssec:g_0_saturation}), inductive
base cases (Subsection ~\ref{ssec:g_0_base}), and exceptional
cases (Subsection ~\ref{ssec:g-0-exceptional}) when $\delta = 0$
before proving the main theorem in full for all $\delta$ in
Subsection ~\ref{ssec:g_0_main}.

@@

\ssec{Exceptional Cases}
\label{ssec:g-0-exceptional}
First, we describe the cases that would not be covered by induction.
Here we present the explicit generators and relations of the remaining
cases given by signatures in the finite set

@@

\noindent
is generated as a $k$-algebra by elements of degree at most $e =
\max(5, e_1, \ldots, e_r)$ and has relations in degree at most $2e$,
so long as $\sigma$ does not lie in a finite list of exceptional
cases, as listed in Table ~\ref{table:g-0-exceptional}.
\end{thm}

@@

Now suppose $\delta = 0$, so $\halfcan$ is not necessarily effective.
Let $\sigma = (0; e_1, \ldots, e_r; 0)$ be a signature with $r > 0$
and not one of the exceptional cases in Table
~\ref{table:g-0-exceptional}.
